---
title: "NESP Kowari project"
output: html_notebook
---
```{r,message=FALSE,include=FALSE}
source("scripts/kowari_script.R")
source("scripts/kowari_habitat_script.R")
```

*Aaron Greenville*  


Kowari capture data from grids: PAN GRID and WAL GRID (approx. 25km apart).  

**To do:**  

* map  

  
##Trapping effort per trip
```{r}
Trap.effort
```

##Trapping effort per year
```{r}
trap.effort.yr
```

##MARSS 1-state model with density dependence
```{r}
kowariMARSS
```

```{r}
caterplot(kowariMARSS, "x")
caterplot(kowariMARSS, "A")
caterplot(kowariMARSS, parms = c("U", "B","sigmaQ", "sigmaR"),
          style= "gray")

```


##MARSS 2-state model with density dependence
```{r}
kowariMARSS.2
```

```{r}
caterplot(kowariMARSS.2, "x")
caterplot(kowariMARSS.2, parms = c("U", "B","sigmaQ", "sigmaR"),
            style= "gray")  
```


##Figures
```{r, echo=FALSE}
par(mfrow = c(1,1), xpd=F, mar=c(5.1, 4.1, 4.1, 11), family = "serif")

matplot(years, (kow.x.ZC.1.TN.r), ylim = c(0,15), type="l",lwd=1, xlab="Year", ylab="Captures (100 trap nights)",
        bty="l", font.lab=2)
  polygon(c(years,rev(years)),c(t(kow.x.ZC.1.TN.LC.r), rev( t(kow.x.ZC.1.TN.UC.r))) ,
        col = adjustcolor("grey90", 0.9), border = NA)
  matlines(years, (kow.x.ZC.1.TN.r), lwd=1)
  matpoints(years,t(kowari.l),pch=15:23, cex =1, col=1)
  legend("topright",xpd=T, legend=sites, pch=15:23, cex=1, pt.cex = 1,
       col=c(1,1),box.col=NA,inset=c(-.33,0))
```
Fig. One-state model.

```{r, echo=FALSE}
par(mfrow = c(1,1), xpd=F, mar=c(5.1, 4.1, 4.1, 11), family = "serif")

matplot(years, (kow.x.2.TN.r), ylim = c(0,15), type="l",lwd=1, xlab="Year", ylab="Captures (100 trap nights)",
        bty="l", font.lab=2)
  polygon(c(years,rev(years)),c(t(kow.x.2.TN.LC.r), rev( t(kow.x.2.TN.UC.r))) ,
        col = adjustcolor("grey90", 0.9), border = NA)
  polygon(c(years,rev(years)),c(t(kow.x.2.TN.2.LC.r), rev( t(kow.x.2.TN.2.UC.r))) ,
        col = adjustcolor("blue", 0.2), border = NA)

  matlines(years, (kow.x.2.TN.r), lwd=2)
  matlines(years, (kow.x.2.TN.2.r), lwd=2, col="blue")
  matpoints(years,t(kowari.l),pch=15:23, cex =1, col=c("black", "blue"))

  legend("topright",xpd=T, legend=sites, pch=15:23, cex=1, pt.cex = 1,
       col=c("black","blue"),box.col=NA,inset=c(-.33,0))
```

Fig. 2-state model.

## Habitat analysis  

nMDS on habitat variables measured around traps on both grids.  

Habitat_LU_cover  
**cover code cover description**  
 1	< 5%  
 2	5 - 25%  
 3	25 - 50%  
 4	50 - 75%  
 5	> 75%  
 
 Gibber Size  	
 1	pebble <50mm  
 2	cobble 50-250mm  
 3	boulder >250mm  

**Habitat variables**
```{r}
colnames(kowari.habitat.m)
```

#### NMDS
```{r}
NMDS.hab
```

```{r}
ggplot.MDS.fn(NMDS.hab, kowari.factor.df, kowari.factor.df$present)
```




